---
title: "O'Grady et al 2019 JGR oceans Extreme Water level estimates"
author: "Julian O'Grady"
date: "2024-03-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
require(terra)

outdir = "plots/"
colz = rev(c("#000078","#0000BB","#0000FF","#0C3AFC","#1975FA","#25B0F7","#32EBF5","#73EEE6","#B4F1D7","#F5F5C8",
             "#F5F564","#F5E032","#F6CC00","#F7A300","#F97A00","#FB5100","#FF0000","#DA0006","#B5000D","#910014",
             "#7B003D","#660066","#88008C","#AB00B2","#CD00D8","#F000FF","#CAA6DE","#DBA7DD","#EDA9D2"))#,"#FFABC7") )


ncolz = length(colz)
colzRamp = colorRampPalette(rev(colz))



```

```{r}

roms = vect("ROMS_gumbel_parameters_SWLandMTWL_Australia")

tg = vect("gesla3_fgumbel_shp")
tg = tg[tg$nCompleteY >= 20]

#Align the ROMs 1 year ARI (63% AEP) location parameter to the nearest tide gauge value following O'Grady et al 2019 and Haigh et al. 2014.
crs(roms) = crs(tg)
nb = nearby(roms,tg)
roms$muOff = tg$mle_1[nb[,2]]

```

```{r}

#predict return levels
AEP = 0.01
return_level <- function(mu,lam,AEP){
  z = mu+lam*(-log(-log(1-AEP)))
  return(z)
}

#storm tide return level
roms$SWL_RL = return_level(roms$muOff,roms$SWLlam,AEP)
#storm tide plus wave setup
roms$MTWL_RL = return_level(roms$muOff-roms$SWLmu+roms$MTWLmu,roms$MTWLlam,AEP)

tg$SWL_RL = return_level(tg$mle_1,tg$mle_2,AEP)

#compute MTWL
```

```{r}

ats = seq(0,7,0.5)
plotit = function(){
  plot(roms,"SWL_RL",breaks=ats,range = range(ats), main = paste0("1% AEP extreme SWL (reletive to MSL of that year)"),type = "continuous",col = colzRamp(length(ats)-1))
  plot(tg,"SWL_RL",add = TRUE,legend=FALSE,col = colzRamp(length(ats)-1),cex = 1.5,breaks=ats,range = range(ats))
  plot(tg,cex = 1.5,pch=1,add=TRUE)
}

plotit()
pdf(paste0(outdir,"One_Percent_AEP_SLW.pdf"),width=12)
plotit()
dev.off()
```
Figure : Height (in m relative to mean sea level in a given year) of the 1% annual exceedance probability (AEP) extreme still water level (SWL) storm tide  (Source: [O’Grady et al 2019](https://agupubs.onlinelibrary.wiley.com/doi/full/10.1029/2018JC014871)) overlayed here with GESLA3 tide gauge record estimates (Source [Canute 3](https://shiny.csiro.au/Canute3_0/)). This shows the height of a water level from the combination of storm surge and tide with a 1% chance of occurring, these heights can be offset with a mean sea level reference to a given year and vertical datum (e.g. AHD) and can be generated for any SLR or AEP for NCRA. Heights can also be generated with the inclusion of damaging wave effect (wave setup [O’Grady et al 2019](https://agupubs.onlinelibrary.wiley.com/doi/full/10.1029/2018JC014871)) for sandy beach coastlines. Here, levelled heights can be integrated with land height data to map coastal flooding. Similar plots based on SCHISM data will be developed by ACS by June 2025.

```{r}

ats = seq(0,7,0.5)
plotit = function() plot(roms,"MTWL_RL",breaks=ats,range = range(ats), main = paste0("1% AEP extreme MTWL (reletive to MSL of that year)"),type = "continuous",col = colzRamp(length(ats)-1))

plotit()
pdf(paste0(outdir,"One_Percent_AEP_MTWL.pdf"),width=12)
plotit()
dev.off()

```
Figure : Height (in m relative to mean sea level in a given year) of the 1% annual exceedance probability (AEP) extreme mean total water level (MTWL) storm tide plus wave setup  (Source: [O’Grady et al 2019](https://agupubs.onlinelibrary.wiley.com/doi/full/10.1029/2018JC014871), Canute 3). This shows the height of a water level from the combination of storm surge and tide and wave setup for sandy beaches with a 1% chance of occurring, these heights can be offset with a mean sea level reference to a given year and vertical datum (e.g. AHD) and can be generated for any SLR or AEP for NCRA. Heights can also be generated without the inclusion of damaging wave effect (still water level O’Grady et al 2009). Here, levelled heights can be integrated with land height data to map coastal flooding. Similar plots based on SCHISM data will be developed by ACS by June 2025.

```{r}
# Set initial Sea Level Rise value
SLR <- 0.6

# Function to plot multiplication factors
plotMF <- function(v, att, ats, buf = 1, SLR = 0.6, ...) {
  v$SLR <- SLR
  
  if ("SWLlam" %in% names(v)) {
    v$SWL_MFSLR_50 <- exp(SLR / v$SWLlam)
    v$SWL_MFSLR_10 <- exp(SLR / (v$SWLlam + 1.28 * v$SWLlam_se))
    v$SWL_MFSLR_90 <- exp(SLR / (v$SWLlam - 1.28 * v$SWLlam_se))
    
    v$MTWL_MFSLR_50 <- exp(SLR / v$MTWLlam)
    v$MTWL_MFSLR_10 <- exp(SLR / (v$MTWLlam + 1.28 * v$MTWLlam_se))
    v$MTWL_MFSLR_90 <- exp(SLR / (v$MTWLlam - 1.28 * v$MTWLlam_se))
  }
  
  if ("mle_2" %in% names(v)) {
    se = sqrt(tg$cov_22)
    v$SWL_MFSLR_50 <- exp(SLR / tg$mle_2)
    v$SWL_MFSLR_10 <- exp(SLR / tg$mle_2 + 1.28 * se)
    v$SWL_MFSLR_90 <- exp(SLR / tg$mle_2 - 1.28 * se)
    
  }
  
  logats <- log(ats)
  
  extu <- as.numeric(as.vector(ext(v)))
  dat <- cut(as.numeric(unlist(values(v[att]))), breaks = ats)
  v$plot <- dat
  levs <- levels(dat)
  
  dumy <- terra::vect(array(0, dim = c(length(ats), 2)), "points")
  dumy$plot <- cut(ats, breaks = c(ats))
  dumy <- rbind(dumy, v)
  dumy <- dumy[order(dumy$plot, decreasing = TRUE)]
  
  terra::plot(dumy, "plot", levels = levs, sort = FALSE, 
              col = rev(viridis::viridis(length(ats - 1))),
              xlim = extu[1:2] + c(-buf, buf), ylim = extu[3:4] + c(-buf, buf), ...)
  
  return(v)
}

# Function to create plots for specific SLR values
plotit <- function(SLR) {
  par(mfrow = c(1, 2))
  
  dat <- plotMF(v = roms, att = "SWL_MFSLR_50", ats = ats, SLR = SLR, 
                main = paste0("Extreme SWL multiplication factor for ", SLR, "m of SLR"), 
                mar = c(2, 2, 2, 4) + 0.1)
  
  ge <- geom(dat)
  
  out <- data.frame(
    longitude = ge[, 3], 
    latitude = ge[, 4], 
    SLR = dat$SLR, 
    SWL_AEP1PC = round(dat$SWL_RL + dat$SLR, 2), 
    MTWL_AEP1PC = round(dat$MTWL_RL + dat$SLR, 2), 
    SWL_MFSLR_50 = round(dat$SWL_MFSLR_50, 2), 
    SWL_MFSLR_10 = round(dat$SWL_MFSLR_10, 2), 
    SWL_MFSLR_90 = round(dat$SWL_MFSLR_90, 2), 
    MTWL_MFSLR_50 = round(dat$MTWL_MFSLR_50, 2), 
    MTWL_MFSLR_10 = round(dat$MTWL_MFSLR_10, 2), 
    MTWL_MFSLR_90 = round(dat$MTWL_MFSLR_90, 2)
  )
  write.csv(out, file = paste0(outdir, "../data/MFSLR_ACSSH_MMALL_SLR", SLR * 100, "cm.csv"), row.names = FALSE)
  
  dattg <- plotMF(v=tg, att="SWL_MFSLR_50", ats = ats, SLR = SLR, cex = 1.5, add = TRUE, legend = FALSE)
  plot(tg, cex = 1.5, SLR = SLR, pch = 1, add = TRUE)
  plotMF(v = roms, att = "MTWL_MFSLR_50", ats = ats, SLR = SLR, 
         main = paste0("Extreme MTWL multiplication factor for ", SLR, "m of SLR"), 
         mar = c(2, 2, 2, 4) + 0.1)

  ge <- geom(dattg)
  dattg <- values(dattg)
  outtg <- data.frame(
    name <- dattg$Name,
    nCompleteY <- dattg$nCompleteY,
    longitude = ge[, 3], 
    latitude = ge[, 4], 
    SLR = dattg$SLR, 
    SWL_AEP1PC = round(dattg$SWL_RL + dattg$SLR, 2), 
    SWL_MFSLR_50 = round(dattg$SWL_MFSLR_50, 2), 
    SWL_MFSLR_10 = round(dattg$SWL_MFSLR_10, 2), 
    SWL_MFSLR_90 = round(dattg$SWL_MFSLR_90, 2) 
  )
  write.csv(outtg, file = paste0(outdir, "../data/MFSLR_GESLA3_MMALL_SLR", SLR * 100, "cm.csv"), row.names = FALSE)
    
}

# Break points for the plots
ats <- c(0, 1, 2, 5, 10, 20, 50, 100, 200, Inf)

# Different SLR values to plot
SLRs <- c(0.06, 0.1, 0.2, 0.38, 0.6, 1.0)
plotit(SLR = 0.6)

# Loop over SLR values and generate plots and CSV files
for (SLR in SLRs) {
  pdf(paste0(outdir, "MFSLR_ACSSH_MM50_SLR_", SLR * 100, "cm.pdf"), width = 14, height = 5)
  plotit(SLR)
  dev.off()
}


```
Figure : Multiplication/amplification factors ([Hunter 2012](https://link.springer.com/article/10.1007/s10584-011-0332-1)) of extreme storm tide water level and mean total water level (storm tide plus wave setup) AEPs (e.g. 1% AEP in Figure A1) for an amount of SLR (see figure title), overlayed here with GESLA3 tide gauge record estimates (Source Canute 3). Here, a AEP with an associated return level is mutiplied by these factors (e.g., 1% x 10) to yeild a new AEP (i.e., 10%) for the associated return level when considering sea level rise. These return levels can be aligned with coastal structure design levels (or DEMs) to understand the change in frequency of exceedence and can be generated for any SLR or AEP for the NCRA.
